\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{array}
\renewcommand{\arraystretch}{1.2}

\title{Bookie Dookie Testing Report}
\author{Senior QA Engineer}
\date{April 29, 2026}

\begin{document}
\begin{titlepage}
\centering
{\Large Lab 2 Task: Automated Software Testing and Analysis}\par
\vspace{0.8cm}
{\Large QA Test Report: Bookie Dookie Web Application}\par
\vspace{1.2cm}
{\large Student Name: Abdelrahman Abdelhafez}\par
\vspace{0.4cm}
{\large Professor Name: Gintarė Kaminskienė}\par
\vspace{0.8cm}
{\large April 28, 2026}\par
\vfill
\thispagestyle{empty}
\end{titlepage}

\tableofcontents
\newpage

\section{Introduction}
This report presents a professional QA automation submission for a full-stack web application. Automated testing is used to ensure functional correctness, regression safety, and consistent behavior across backend APIs and frontend logic.

\section{Testing Framework Justification}
\textbf{Pytest + pytest-django} were selected for Django compatibility, fixture-driven test isolation, and scalable reporting. The framework supports database rollbacks and clean Arrange-Act-Assert (AAA) structure.\\
\textbf{Jest} was selected for fast execution of JavaScript logic tests with isolated DOM emulation, making it suitable for unit-level frontend behavior.

\section{Mapping to Requirements}
\begin{itemize}
    \item \textbf{Unit Testing}: CRUD operations and model constraints (create, update, delete, boundary checks).
    \item \textbf{Integration Testing}: API list/detail endpoints and response schema consistency. Search behavior is validated on the frontend because the backend provides list retrieval only.
    \item \textbf{Error Handling}: Missing/invalid inputs, non-existent IDs, and large payload behavior.
\end{itemize}

\section{Test Design}
All tests follow the AAA pattern with explicit Arrange, Act, and Assert steps. Isolation is enforced via pytest-django database fixtures and APIClient usage. Each test is independent and deterministic.

\section{Project File Structure (Tests)}
The workspace is split into backend and frontend directories. Test cases are maintained in the following files:
\begin{itemize}
    \item \textbf{Backend tests (pytest)}: \texttt{Bookie\_Dookie/Bookie\_Dookie\_Dj/tests/test\_unit.py}, \texttt{test\_integration.py}, \texttt{test\_errors.py}.
    \item \textbf{Backend fixtures}: \texttt{Bookie\_Dookie/Bookie\_Dookie\_Dj/tests/conftest.py} defines shared data and authenticated clients.
    \item \textbf{Frontend tests (Jest)}: \texttt{Bookie-Dookie/tests/book\_features.test.js} and \texttt{admin.test.js}.
    \item \textbf{Test runner}: \texttt{run\_tests.py} controls execution and logging.
    \item \textbf{Test configuration}: \texttt{Bookie\_Dookie/Bookie\_Dookie\_Dj/pytest.ini}, \texttt{Bookie-Dookie/jest.config.js}, and \texttt{Bookie-Dookie/package.json}.
\end{itemize}
To add or modify test cases, edit the relevant backend or frontend test files listed above. Fixtures should be updated only when shared setup needs to change.

\section{Test Cases (Detailed)}
\textbf{Backend Unit Tests}
\begin{itemize}
    \item Create book with valid data and verify persisted fields.
    \item Create duplicate books with case/spacing variations.
    \item Boundary validation at maximum field lengths (accept 100/200, reject 101+).
    \item Special characters accepted in title/author/category.
    \item Partial update rejected when required fields are missing.
    \item Delete existing book and verify removal.
\end{itemize}

\textbf{Backend Integration Tests}
\begin{itemize}
    \item Retrieve all books and validate list length.
    \item Retrieve single book by ID.
    \item Validate response schema for list endpoint (field consistency).
    \item Dashboard user list retrieval.
    \item Borrowing list filtered by authenticated user.
    \item Empty book\_id query returns full list.
\end{itemize}

\textbf{Backend Error Handling}
\begin{itemize}
    \item Missing multiple required fields in create request.
    \item Invalid book ID (non-numeric) and whitespace-only ID (known behavior).
    \item Non-existent IDs for delete and update.
    \item Extremely large description payload.
    \item Whitespace-only inputs rejected.
\end{itemize}

\textbf{Frontend Logic Tests}
\begin{itemize}
    \item Search is case-insensitive.
    \item Empty query returns all items.
    \item No match returns empty results.
    \item Pagination: first/last page button states.
    \item Pagination: empty dataset handling (documented current behavior).
\end{itemize}

\section{Edge Case Analysis}
\begin{itemize}
    \item \textbf{Overlong strings}: validate maximum field constraints to prevent DB or UI overflow.
    \item \textbf{Whitespace-only inputs}: prevent silent acceptance of unusable values.
    \item \textbf{Non-numeric IDs}: reveal API robustness for invalid query parameters.
    \item \textbf{Large payloads}: confirm stability with higher-than-average inputs.
    \item \textbf{Empty dataset pagination}: ensures UI state remains predictable when no records exist.
\end{itemize}

\section{Test Case Matrix}
The following tables document each automated test case, including input values, expected outcome, actual outcome, and the testing module used.

\subsection{Backend Test Cases (pytest)}
\begin{longtable}{|>{\raggedright\arraybackslash}p{2.4cm}|>{\raggedright\arraybackslash}p{1.8cm}|>{\raggedright\arraybackslash}p{3.6cm}|>{\raggedright\arraybackslash}p{2.6cm}|>{\raggedright\arraybackslash}p{2.0cm}|>{\raggedright\arraybackslash}p{1.4cm}|>{\raggedright\arraybackslash}p{1.2cm}|}
\hline
\textbf{Test Case} & \textbf{Module} & \textbf{Input Values} & \textbf{Expected Result} & \textbf{Actual Result} & \textbf{Tool} & \textbf{Status} \\ \hline
\endfirsthead
\hline
\textbf{Test Case} & \textbf{Module} & \textbf{Input Values} & \textbf{Expected Result} & \textbf{Actual Result} & \textbf{Tool} & \textbf{Status} \\ \hline
\endhead
Add book missing fields & Books API & title, author, cover\_url only & 400 validation error & 400 (validation errors) & pytest & PASS \\ \hline
Delete non-existent book & Books API & book\_id=9999 & 404 not found & 404 Not Found & pytest & PASS \\ \hline
Borrowings empty list & Users API & no borrowed records & empty list & [] (length=0) & pytest & PASS \\ \hline
Get non-existent book & Books API & book\_id=9999 & empty serializer fields & 200 with empty fields & pytest & PASS \\ \hline
Missing multiple fields & Books API & title only & 400 validation error & 400 (missing fields) & pytest & PASS \\ \hline
Invalid ID delete (current) & Books API & book\_id=not-an-int & raises ValueError & ValueError thrown & pytest & PASS \\ \hline
Edit non-existent ID & Books API & book\_id=9999 + full payload & 404 not found & 404 Not Found & pytest & PASS \\ \hline
Delete already deleted & Books API & delete same ID twice & 404 on second delete & 404 Not Found & pytest & PASS \\ \hline
Invalid ID expected 400 & Books API & book\_id=not-an-int & 400 expected & ValueError thrown & pytest & FAIL \\ \hline
Whitespace ID expected list & Books API & book\_id=" " & list of books & ValueError thrown & pytest & FAIL \\ \hline
Get book list & Books API & no params & list length=2 & 200 list (length=2) & pytest & PASS \\ \hline
Get book by ID & Books API & book\_id=valid & correct book data & 200 with matching fields & pytest & PASS \\ \hline
Dashboard users list & Dashboard API & none & list of users & 200 list (>=2 users) & pytest & PASS \\ \hline
Borrowings filtered & Users API & user has 1 borrowing & list length=1 & 200 list (length=1) & pytest & PASS \\ \hline
Empty book\_id query & Books API & book\_id="" & full list returned & 200 list (length=2) & pytest & PASS \\ \hline
Response schema fields & Books API & list endpoint & fixed field set & keys match schema & pytest & PASS \\ \hline
Whitespace query ValueError & Books API & book\_id=" " & ValueError raised & ValueError thrown & pytest & PASS \\ \hline
Add book valid data & Books API & normal payload & 201 created & 201 Created & pytest & PASS \\ \hline
Edit book full update & Books API & full payload & 200 updated & 200 Updated & pytest & PASS \\ \hline
Delete book success & Books API & valid ID & 200 deleted & 200 Deleted & pytest & PASS \\ \hline
Duplicate username & Users API & same username twice & 400 on second & 400 duplicate username & pytest & PASS \\ \hline
Duplicate books case/space & Books API & "Dune" + "  DUNE " & both saved & 201 + 201 created & pytest & PASS \\ \hline
Overlong fields rejected & Books API & 101/201 chars & 400 validation error & 400 (max length) & pytest & PASS \\ \hline
Max length accepted & Books API & 100/200 chars & 201 created & 201 Created & pytest & PASS \\ \hline
Special characters & Books API & symbols in fields & 201 created & 201 Created & pytest & PASS \\ \hline
Whitespace-only inputs & Books API & spaces in fields & 400 validation error & 400 (blank values) & pytest & PASS \\ \hline
Partial update rejected & Books API & title only & 400 validation error & 400 (required fields) & pytest & PASS \\ \hline
Large description & Books API & description 5000 chars & 201 created & 201 Created & pytest & PASS \\ \hline
\end{longtable}

\subsection{Frontend Test Cases (Jest)}
\begin{longtable}{|>{\raggedright\arraybackslash}p{2.8cm}|>{\raggedright\arraybackslash}p{1.8cm}|>{\raggedright\arraybackslash}p{3.6cm}|>{\raggedright\arraybackslash}p{2.4cm}|>{\raggedright\arraybackslash}p{1.6cm}|>{\raggedright\arraybackslash}p{1.0cm}|>{\raggedright\arraybackslash}p{1.0cm}|}
\hline
\textbf{Test Case} & \textbf{Module} & \textbf{Input Values} & \textbf{Expected Result} & \textbf{Actual Result} & \textbf{Tool} & \textbf{Status} \\ \hline
\endfirsthead
\hline
\textbf{Test Case} & \textbf{Module} & \textbf{Input Values} & \textbf{Expected Result} & \textbf{Actual Result} & \textbf{Tool} & \textbf{Status} \\ \hline
\endhead
Search filters fields & book\_features.js & query="clean" & only matching book & 1 match returned & Jest & PASS \\ \hline
Search no match & book\_features.js & query="missing" & empty list & [] returned & Jest & PASS \\ \hline
Search empty query & book\_features.js & query="" & all books returned & full list returned & Jest & PASS \\ \hline
Search case-insensitive & book\_features.js & query="dUnE" & matches Dune & Dune returned & Jest & PASS \\ \hline
Pagination base state & admin.js & totalUsers=10, page=1 & Page 1 of 1, buttons disabled & prev=true, next=true & Jest & PASS \\ \hline
Pagination empty dataset & admin.js & totalUsers=0, page=1 & Page 1 of 0, next disabled & next=false (known issue) & Jest & FAIL \\ \hline
Pagination last page & admin.js & totalUsers=30, page=3 & Page 3 of 3, next disabled & prev=false, next=true & Jest & PASS \\ \hline
\end{longtable}

\section{Test Execution Procedure}
\begin{enumerate}
    \item Install backend dependencies: \texttt{pip install -r requirements.txt}.
    \item Install backend test dependencies: \texttt{pip install -r requirements-test.txt}.
    \item Install frontend dependencies: \texttt{npm install} (portable Node.js in \texttt{.tools/node}).
    \item Run backend tests (verbose): \texttt{pytest -vv}.
    \item Run backend coverage: \texttt{pytest -vv --cov=Books --cov=Users --cov=Dashboard --cov-report=term-missing}.
    \item Run frontend tests with JSON output: \texttt{npm test -- --runInBand --json --outputFile=jest-results.json}.
    \item Optional: run \texttt{python run\_tests.py} for step logging to \texttt{test\_run.log}.
    \item Run known-issue backend tests only: \texttt{pytest -vv -m known\_issue}.
    \item Run known-issue frontend tests only: \texttt{npm test -- --runInBand --testPathPattern=known\_issues.test.js --testPathIgnorePatterns=}.
\end{enumerate}

\section{Test Execution Results}
\begin{itemize}
    \item Backend baseline run: 26 passed, 2 xfailed, 0 failed
    \item Backend known-issue run: 0 passed, 2 failed
    \item Frontend baseline run: 7 passed, 0 failed
    \item Frontend known-issue run: 0 passed, 1 failed
\end{itemize}

\section{Execution Output (Evidence)}
\textbf{Pytest Summary}\
\begin{verbatim}
26 passed, 2 xfailed in 9.42s
\end{verbatim}

\textbf{Pytest Known-Issue Summary}\
\begin{verbatim}
2 failed, 28 deselected in 1.20s
\end{verbatim}

\textbf{Coverage Summary}\
\begin{verbatim}
TOTAL                                                             266     53    80%
\end{verbatim}

\textbf{Jest Summary}\
\begin{verbatim}
Test Suites: 2 passed, 2 total
Tests:       7 passed, 7 total
\end{verbatim}

\textbf{Jest Known-Issue Summary}\
\begin{verbatim}
Test Suites: 1 failed, 1 total
Tests:       1 failed, 1 total
\end{verbatim}

\section{Coverage Analysis}
Total backend coverage is \textbf{80\%}. Gaps are concentrated in authentication-oriented user views and unused default test files. Core CRUD and list paths are fully exercised.

\section{Issues Encountered and Fixes}
\begin{itemize}
    \item \textbf{Invalid book\_id:} backend raises \texttt{ValueError} for non-numeric IDs; documented as \texttt{xfail} to reflect expected failure.
    \item \textbf{Whitespace book\_id:} same behavior documented as \texttt{xfail}.
    \item \textbf{Empty dataset pagination:} current UI allows next button; test asserts observed behavior and reports as a known UI limitation.
\end{itemize}

\section{System Reliability Assessment}
The system demonstrates stable CRUD and list behaviors with strong backend coverage and validated frontend search/pagination logic. Known issues are constrained to input validation robustness and UI pagination edge cases, which are documented for future remediation.

\section{How to Run Tests}
Run the menu script from the workspace root:
\begin{verbatim}
python run_tests.py
\end{verbatim}
Backend tests directly:
\begin{verbatim}
cd Bookie_Dookie/Bookie_Dookie_Dj
pytest -vv --cov=Books --cov=Users --cov=Dashboard --cov-report=term-missing
\end{verbatim}
Frontend tests:
\begin{verbatim}
cd Bookie-Dookie
npm test
\end{verbatim}

\end{document}
